### 2025-04-30 docs/Homeownership_BankOfMomAndDad.tex
Done with empirical section

\textit{AHS:} I use the 1991-2019 AHS, which surveys occupants of housing units, to construct a time series of whether homeowners primarily funded their down payment with gifts or inheritances (right panel of Figure~\ref{fig:motivation}). After excluding households who report sale of previous homes as a source, the sample includes about 50,000 observations each year. The series is relatively flat through the 1990s, followed by a sharp decline from 2005 to 2009---coinciding with the rise of low–down-payment mortgages---and a marked increase from 2013 onward. Because the horizontal axis reflects survey year rather than purchase year, minor fluctuations largely reflect variation among recent buyers, who make up a small share of all homeowners.

@@

I use data from 1999 to 2021. In 1999, the PSID started to collect detailed wealth data every other year. In most waves of the PSID, there is limited transfer data, and the main question is whether households received gifts or bequests over \$5,000 in the last two years. In 2013, the PSID collected more detailed transfer data in the Family Roster and Transfer Module. They asked parents how much they gave their children in the last calendar year and how much they had given over their lifetime for school, house purchases, or other purposes. Household characteristics such as age, gender, and education refer to the household head. I classify top-coded values as missing observations. All monetary variables are expressed in thousands of 2016 U.S. dollars. Table \ref{tab:vardef} in the Appendix provides variable code of the main variables.

@@

\textit{Matching Parents with Children:} I use the Parent/Child file from the PSID's 2013 transfer supplement. I can observe each household's reported transfers to and from parents and children. This leads to discrepancies, where the child and parent do not agree on the amount given from the parent to the child. I only use the parent's reported transfer. First, there may be some stigma about receiving transfers, which may induce receiving children to underreport. Second, in the model, parents determine the size of the transfer they give to their child.

\textit{Definition of Transfers:} To measure the transfer receipt rate, I rely on the transfer supplement which asked all parent households whether they gave money, gifts, or loans of \$100 or more to each children in 2012. I follow the literature \citep[e.g.,][]{mcgarry2016dynamic} and treat all transfers as gifts. Since this paper focuses on transfers that (a) relate to housing and (b) are quantitatively meaningful, I set transfers below \$500 to \$0. About 25\% of transfers are below this threshold, and ignoring them increases the conditional mean transfer from \$2,921 to \$3,960 and reduce the transfer receipt rate from 45\% to 22\%.\footnote{These transfer rates are relatively high and higher than those calculated by \citealt{feiveson2019lifecycle} using the Surey of Consumer Finances (SCF). First, it seems likely that the PSID transfer supplement was more successfull in counting transfers than the SCF. Second, the SCF asks about transfers and gifts going further back in time, likely increasing underreporting. \textcolor{red}{write more here} }

In all other waves, there is a general question of gifts and inhertitances from people outside the household. We do not know the source of these transfers (e.g., they could be from friends or an inheritance from grandparents) and hence these do not know the extent to which these captures parental assistance. Yet, they, to some extent, capture the impact of transfer receipt on household decisions, and hence I use these variables in some of the empirical results, following \cite{Lee2018}.

@@

\subsubsection{Descriptive Statistics: Who Receives Transfers?}
\textcolor{red}{!! [[rewrite this]] !!}
I now discuss descriptive statistics from the PSID sample in 2013 for households aged 25 to 44 with an observed parent household. Table \ref{tab:descrstats} contains the means of variables, by age, wealth, homeownership, and transfer receipt. Though limited to households with observed parents, the sample appears reasonably representative. Average income by age closely matches estimates from the Survey of Consumer Finances \cite[see e.g.,][]{kuhn20162013}, though average wealth is slightly lower.

There are several main takeaways from the subgroup analysis. In 2013, 22\% of young households received a transfer, and transfers averaged \$3,950. Receivers have significantly richer parents, have similar wealth and income as nonreceivers, and are less likely to own. Receivers are more likely to transition from renting to owning, especially in the age groups where households are most likely to buy. The key determinant of transfer receipt seems to be parental wealth.

In panel a), the first two columns compare the whole sample, by transfers receipt in the last calendar year. The mean transfer is relatively high at \$3,950, and 22\% of households received transfers in the last year. The largest difference is parental wealth: transferring parents are 2.5 times as wealthy as nontransferring parents (\$1,029,000 vs. \$410,000). Perhaps surprisingly, receivers are slightly richer than nonreceivers (\$108,000 vs. \$78,500), are more likely to be college-educated and white, and are one year younger. Receivers are slightly less likely to be homeowners (39\% vs. 43\%), reflecting the age difference and the tendency for college attendance to delay homeownership.

\begin{table}[htbp]
	\begin{adjustbox}{center}
		\small
		\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
		\begin{threeparttable}
			\singlespacing
			\caption{Descriptive Statistics (Means), Households Aged 25-44}\label{tab:descrstats}
			\begin{tabular}{l  r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r}
				\toprule
				& \multicolumn{2}{c}{All}  &  \multicolumn{2}{c}{Renter}   & \multicolumn{2}{c}{Owner}    \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
				Receiver 		& \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\  \midrule
				\input ../tabfig/descr/descr_trsnf_disc_mean.tex
				\midrule
				Age                    & \multicolumn{2}{c}{25-31} & \multicolumn{2}{c}{32-38}  & \multicolumn{2}{c}{39-44}   \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\ 
				 \midrule
				\input ../tabfig/descr/descr_trsnf_age_mean.tex
				\midrule
				Wealth Tertile                    & \multicolumn{2}{c}{Tertile 1}  & \multicolumn{2}{c}{Tertile 2}  & \multicolumn{2}{c}{Tertile 3}    \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\ 
				\midrule
				\input ../tabfig/descr/descr_trsnf_wlt_mean.tex
		\bottomrule
			\end{tabular}
		\end{threeparttable}
	\end{adjustbox}
	{\begin{footnotesize}\begin{flushleft}\vspace{-0.1in}%
		\textit{Notes:} Data from the PSID Transfer, Individual, and Family modules. Weighted using family weights. Transfer, wealth, and income measured in 1000s of 2016 USD. 
	\end{flushleft}\end{footnotesize}}		
	\end{table}

Next, I break the sample down by ownership. Homeowners have 10 times the wealth of renters and also significantly wealthier parents. Renters and owners receive transfers at about the same rate and size. Receivers are more likely to switch from renting to owning: 21\% of receiving owners rented two years ago, compared to 14\% of nonreceiving owners.

@@

While a growing literature documents that parental transfers increase the probability that renters become homeowners (e.g.\ \citealp{wold2024housing}; \citealp{Blickle2019}; \citealp{benetton2022dynastic}), I replicate this result to highlight the importance of transfers for housing outcomes. I estimate
\begin{equation}	\label{eq:transferrent}
	\Pr\left(Own_{i,t+2}=1 \mid Own_{i,t}=0\right)
	= \alpha + \beta\,Transfer_{i,t+2}
	  + \gamma' X_{i,t} + \varepsilon_{i,t},
\end{equation}
where \(Own_{i,t}=0\) denotes a renter at time \(t\), \(Transfer_{i,t}=1\) if the household received a transfer between survey waves, and \(X_{i,t}\) includes log parental wealth, log household income and log household net worth, log household size, age, parental age, and dummies for  education, marital status, race, household composition changes, and state and year fixed effects. The sample is limited to renting households aged 25 to 44 who have never been observed as homeowners. Table \ref{tab:newowners_short} reports the estimated coefficient on transfer ($\beta$), with the full coefficient table relegated to Appendix Table \ref{tab:newowners}. These regressions are similar to those in \cite{Lee2018}, who also use the PSID.

To measure transfer receipt, I compare two definitions. Definition 1 (narrow): Parental gifts exceeding \$500 in 2012, as reported in the PSID’s 2013 Transfer Module, limiting the sample to renters in 2011.  Definition 2 (broad): Any gift or inheritance exceeding \$10,000 reported in all waves of the PSID, or a parental gift exceeding \$10,000 in 2012. This expands the sample include all waves of the PSID, and lets us include household fixed effects, though we now only capture large transfers and include some non-parental transfers.

@@

To increase statistical power, I now use the second transfer receipt definition, which also includes non-parental gifts and bequests. This leads to a ten-fold increase in sample size, but the transfer receipt rate also decrease to around 3\%, and the median transfer size jumps from \$1,500 to \$24,200. In addition, we can now follow renters over time, enabling the inclusion of household fixed effect. In a univariate regression, transfers increase the probability of becoming a homeowner by 21.9 percentage points. Even with the full set of controls, the effect of transfer receipt on the future homeownership is large and statisticaly significant: 13.5 percentage point in the OLS regression ($p<0.001$) and a 7.9 percentage point in the FE regression  ($p<0.1$). These effects are large, similar in magnitude of the baseline rent-to-own transition rate of 18 percent. A few caveats are in order: I only capture transfers above \$10,000, and the effect of transfer receipt is likely increasing with transfer size. Second, these transfers are not exogenous: indeed, it is likely that some of these transfers are given \emph{because} households are buying a house.

@@

\subsubsection{Transfer Sizes}
For additional detail on the transfer variables, I plot kernel density estimates in Figure \ref{fig:PSID_transfer_to_renters}, for renters in $t$, separated by whether they own or rent in $t+2$. The left panel only includes parental transfers exceeding \$500 from the 2013 transfer supplement. The right panel only includes transfer/inheritance exceeding \$10,000 which is included in every wave. There are two takeaways. First, households who transition to ownership receive larger transfers on average, under both transfer measures. Second, the transfers in the transfer supplement are much smaller than than those in the gift/inheritance question. Indeed, the median parental transfer is about \$1k, the mean about \$3k, and the 95th percentile is at \$10k, compared to the median gift/inheritance of \$25k, mean of \$60k, and a 95th percentile of about \$130k.

While these are the best data available to follow household over time---in particular their wealth, income, and housing choices---and transfers, the data have several some limitations. First, in the transfer supplement we only observe transfers in 2012, while the PSID is collected every other year. Hence we undercount both transfer rates and sizes. Second, in the gift/inheritance question, we do not know the source of the transfer. Third, we do not know the extent to which households report intra-family loans, nor the extent to which these are repaid.
\begin{figure}
	\caption{Transfer Sizes to Renters by Transition to Ownership}\label{fig:PSID_transfer_to_renters}
	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_transfers_renters_transfers.pdf}% 
	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_transfers_renters_inheritance.pdf}%

@@

\subsubsection{Households with Wealthier Parents Are More Likely to Remain Homeowners}
While the previous sections have focused on the transition into ownership, I now focus on mainting homeownership. \textcolor{red}{rewrite this better: That said, \cite{bond2021role} is an exception and shows the positive association between parental wealth and children's housing outcomes in the PSID and the Health and Retirement Study. }

@@

To understand which variables predicts retainment of ownership, I estimate
\begin{equation}	\label{eq:regmaintain}
	\Pr\left(Own_{i,t+\Delta}=1 \mid Own_{i,t}=1\right)
	= \beta\ln(NetWorth_{i,t}) + \zeta\ln(ParWealth_{i,t})
	  + \gamma' X_{i,t} + \varepsilon_{i,t},
\end{equation}

where $Own_{i,t+\Delta}=1$ denotes an owner at time $t+\Delta$, $\beta$ and $\zeta$ measures the effect of log household net worth and parental wealth, respectively  and $X_{i,t}$ is a vector of control variables. The only changes relative to Eq. \ref{eq:transferrent} is that the outcome variable conditions on current ownership and measures future ownership, that there are no controls for transfer receipt, and that the household net worth and parental wealth are taken out of the vector of control variables. The sample is limited to households aged 25 to 44. These regressions are motivated by \cite{bond2021role}, who estimate similar equations in the PSID and the Health and Retirement Survey. Figures \ref{fig:coef_allowner} and \ref{fig:coef_ftowner} plots the two coefficients of interest, over variying lead, times and their 90\% confidence intervals for all only and first-time owners only.

Two results stand out: First, we see that parental wealth is at least as important as household net worth at all horizons, though the difference is never statistically significant. Second, we see that the impact of parental wealth is increasing over time. For example, we see that a 10\% increase in parental wealth increases the probability of retaining ownership after 8 years by 0.45 percentage points among first-time owners. These patterns will be serve as one part of model validation.
\begin{figure}[tb]
	\captionsetup[subfigure]{aboveskip=-1pt}
	\caption{Mainting Homeownership}\label{fig:homeowner_exit}
	\begin{subfigure}{0.33\textwidth}%
		\caption{Exit Out of Ownership}\label{fig:owner_hazard}%
		\includegraphics[width=\textwidth]{../tabfig/descr/PSID_ownerexit.pdf}% 
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.33\textwidth}%
		\caption{All Owners}\label{fig:coef_allowner}%
		\includegraphics[width=\textwidth]{../tabfig/descr/PSID_coefowner.pdf}% 
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.33\textwidth}%
		\caption{First-Time Owners}\label{fig:coef_ftowner}%
		\includegraphics[width=\textwidth]{../tabfig/descr/PSID_coefftowner.pdf}% 
	\end{subfigure}%
	\hfill

	\caption*{\footnotesize \textit{Notes:} \textcolor{red}{write this footnote.}}
\end{figure}

\subsubsection{Households with Wealthier Parents Are Less Likely to Be Behind on Mortgages} 
% To realize the benefits of homeownership, households must be able to maintain their mortgage payments, since failure to do so can lead to distressed and forced sales. I now show that households with wealthier parents are less likely to be behind on their mortgages, even though they buy more expensive houses. The PSID has collected this data since 2009, and 3.9\% of owners aged 25 to 44 are behind, on average, while 9.2\% have been behind at least once.
One potential roadblock to maintaining ownership are mortgage delinquencies, as failure to meet mortgage payments can lead to distressed and forced sales. I now show that households with wealthier parents are less likely to be behind on their mortgages, even though they buy more expensive houses. The PSID has collected this data since 2009, and about 2\% of owners aged 25 to 44 are behind.

To understand which variables predicts whether households are behind on mortgage payments, I estimate
\begin{equation*}
\Pr(Behind_{i,t+2}) = \beta\ln(NetWorth_{i,t}) + \zeta\ln(ParWealth_{i,t}) + \gamma' X_{i,t} + \varepsilon_{i,t},
\end{equation*}
and the control variables are the same as in the previous regression. Table \ref{tab:hypo} reports the estimated coefficient on household net worth ($\beta$) and parental wealth ($\zeta$), with the full coefficient table relegated to Appendix Table \ref{tab:hypo_long}.

I first limit the sample to first-time buyers to focus on distress among new owners. In a univariate regression, a 10\% increase in parental wealth is associated with a 0.07 percentage point decrease in the probability of being behind on mortgages two years later ($p<0.1$) WWith the full set of controls, the effect increases to 0.1 percentage points, though it is no longer statistically significant. These are economically meaningful since 1.6\% of first-time owners are behind on their mortgage. Moreover, parental wealth is more predictive of future delinqucies than household net worth. In Columns 3 to 5, I instead focus on all owners (who are currently not behind), increasing the sample sizes and allowing us to include household fixed effects. In these specifications, a 10\% increase in parental wealth decreases the probability of being behind two years later by 0.09 percentage points in the univariate OLS regression ($p<0.001$), 0.02 percentage points in the OLS regression (not significant) and  by 0.07 percentage points in the FE regression ($p<0.1$). All told, the probability of being behind on mortgages is decreasing in parental wealth, highlighting that parental resources appear to help new homeowners maintain ownership after purchase---in addition to relaxing borrowing constraints.
\begin{table}
	\centering
	\begin{threeparttable}
		\caption{Housing Choices and Parental Wealth}
		\label{tab:hypo}
		\small 
				\input{../tabfig/regr/hypoII_short.tex}

@@

\subsubsection{Households with Wealthier Parents Are Less Likely to Downsize during Unemployment Spells}\label{sec:eventstudy}
Distressed and forced home sales are costly and frequently driven by unemployment \citep{kermani2021racial,hsu2018unemployment}. If parental wealth provides insurance, households with wealthier parents should be less likely to downsize after income losses. I perform a simple event study on the effect of unemployment on housing consumption to test this hypothesis.

The exercise follows \cite{Chetty2007} closely. The outcome of interest is the change in log housing consumption, which is set to zero for households who do not move. For households who do move, housing consumption is defined as yearly rent while renting and the imputed rent of owner-occupied housing is set to 5\% of the market value \citep{Davis2008}. The sample is limited to household heads who are unemployed only once between ages 25 and 44. I divide the sample by whether a household's parents were in the top parental wealth quartile (of parents with children aged 25 to 44) at the time of unemployment and compare the two groups' housing consumption growth rates at unemployment. The results are displayed in Figure \ref{fig:housinggrowthrates}. Among households with nonwealthy parents, housing consumption decreases by about 5\% during unemployment (significant at the 95\% level). In contrast, among households with wealthy parents, housing consumption even grows slightly at unemployment. Moreover, I replicate this event study using simulated data from the quantitative model and show that the results are similar (see Section \ref{sec:externalval} and Figure \ref{fig:housinggrowthrates_model}). In appendix \ref{app:eventstudy_details} I, instead of using the means, use a formal event study, which shows that households with poor parents have a statsticaly significant decrease in housing consumption during unemployment spells. 
\begin{figure}
	\caption{Event Study: Housing Consumption at Unemployment by Parental Wealth}\label{fig:housinggrowthrates}
	\noindent\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthpoor_both}\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthrich_both}

>>>>

\textit{AHS:} I use the 1991-2019 AHS, which surveys occupants of housing units, to construct a time series of whether homeowners primarily funded their down payment with gifts or inheritances (right panel). After excluding households who report sale of previous homes as a funding source, the sample includes about 50,000 observations each year. The series is relatively flat through the 1990s, followed by a sharp decline from 2005 to 2009---coinciding with the rise of low–down-payment mortgages---and a marked increase from 2013 onward. Because the horizontal axis reflects survey year rather than purchase year, minor fluctuations largely reflect variation among recent buyers, who make up a small share of all homeowners.

@@

I use data from 1999 to 2021. In 1999, the PSID started to collect detailed wealth data every other year. In most waves of the PSID, there is limited transfer data, and the main question is whether households received gifts or bequests over \$5,000 in the last two years. In 2013, the PSID collected more detailed transfer data in the Family Roster and Transfer Module. They asked parents how much they gave their children in the last calendar year and how much they had given over their lifetime for school, house purchases, or other purposes. Household characteristics such as age, gender, and education refer to the household head. I classify top-coded values as missing observations. All monetary variables are expressed in thousands of 2016 U.S. dollars.

@@

\textit{Matching Parents with Children:} I use the Parent/Child file from the PSID's 2013 transfer supplement. I can observe each household's reported transfers to and from parents and children. This leads to discrepancies, where the child and parent do not agree on the amount given from the parent to the child. For each child, I use the parent's reported transfer to that child. First, there may be some stigma about receiving transfers, which may induce receiving children to underreport. Second, in the model, parents determine the size of the transfer they give to their child.

\textit{Definition of Transfers:} To measure the transfer receipt rate, I rely on the transfer supplement which asked all parent households whether they gave money, gifts, or loans of \$100 or more to each child in 2012. I follow the literature \citep[e.g.,][]{mcgarry2016dynamic} and treat all transfers as gifts. Since this paper focuses on transfers that (a) relate to housing and (b) are quantitatively meaningful, I ignore transfers below \$500. About 25\% of transfers are below this threshold, and ignoring them increases the conditional mean transfer from \$2,921 to \$3,960 and reduces the transfer receipt rate from 45\% to 22\%.\footnote{These transfer rates are relatively high and exceed those reported by \citealt{feiveson2019lifecycle} using the Survey of Consumer Finances (SCF). The PSID transfer supplement likely captures a broader range of transfers than the SCF, which deliberately focuses on substantial transfers. In addition, the SCF asks about transfers further back in time, increasing the risk of recall error and underreporting.}

In all other waves, there is a general question of gifts and inhertitances from people outside the household. We do not know the source of these transfers (e.g., they could be from friends or grandparents) and hence these do not know the extent to which these captures parental assistance. Nevertheless, they still capture some of the impact of transfer receipt on household decisions. I therefore include these variables in some of the empirical results, following \citet{Lee2018}.

@@

\subsubsection{Descriptive Statistics: Who Receives Transfers?}
I now discuss descriptive statistics from the 2013 PSID sample for households aged 25 to 44 with an observed parent household. Table \ref{tab:descrstats} contains the means of variables, by age, wealth, homeownership, and transfer receipt. Though limited to households with observed parents, the sample appears reasonably representative. Average income by age closely matches estimates from the SCF \cite[see e.g.,][]{kuhn20162013}, though average wealth is slightly lower.

There are several main takeaways from the subgroup analysis. In 2012, 21\% of young households received a transfer, and transfers averaged \$3,960. Receivers have significantly richer parents, have similar wealth and income as nonreceivers, and are less likely to own. Receivers are more likely to transition from renting to owning, especially in the age groups where households are most likely to buy. The key determinant of transfer receipt seems to be parental wealth.

The first two columns of the top panel compare households by transfers receipt. The mean transfer is relatively high at \$3,960, and 21\% of households received transfers in the last year. The largest difference is parental wealth: transferring parents are 2.5 times as wealthy as nontransferring parents. Perhaps surprisingly, receivers are slightly richer than nonreceivers (\$108,000 vs. \$78,500), are more likely to be college-educated and white, and are one year younger. Receivers are slightly less likely to be homeowners (39\% vs. 43\%), reflecting the age difference and the tendency for college attendance to delay homeownership.

\begin{table}[tbp]
	\begin{adjustbox}{center}
		\small
		\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
		\begin{threeparttable}
			\singlespacing
			\caption{Descriptive Statistics (Means), Households Aged 25-44}\label{tab:descrstats}
			\begin{tabular}{l  r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r}
				\toprule
				& \multicolumn{2}{c}{All}  &  \multicolumn{2}{c}{Renter}   & \multicolumn{2}{c}{Owner}    \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
				Receiver 		& \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\  \midrule
				\input ../tabfig/descr/descr_trsnf_disc_mean.tex
				\midrule
				Age                    & \multicolumn{2}{c}{25-31} & \multicolumn{2}{c}{32-38}  & \multicolumn{2}{c}{39-44}   \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\ 
				 \midrule
				\input ../tabfig/descr/descr_trsnf_age_mean.tex
				\midrule
				Wealth Tertile                    & \multicolumn{2}{c}{Tertile 1}  & \multicolumn{2}{c}{Tertile 2}  & \multicolumn{2}{c}{Tertile 3}    \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\ 
				\midrule
				\input ../tabfig/descr/descr_trsnf_wlt_mean.tex
		\bottomrule
			\end{tabular}
		\end{threeparttable}
	\end{adjustbox}
	{\begin{footnotesize}\begin{flushleft}\vspace{-0.1in}%
		\textit{Notes:} Data from the PSID Transfer, Individual, and Family modules. Weighted using family weights. Transfer, wealth, and income measured in 1000s of 2016 USD. Table \ref{tab:vardef} in the Appendix provides variable codes of the main variables.
	\end{flushleft}\end{footnotesize}}		
	\end{table}

Next, I break the sample down by ownership. Homeowners are both wealthier and have wealthier parents than renters. Renters and owners receive transfers at about the same rate and size. Receivers are more likely to switch from renting to owning: 21\% of receiving owners rented two years ago, compared to 14\% of nonreceiving owners.

@@

While a growing literature documents that parental transfers increase the probability that renters become homeowners (e.g.\ \citealp{wold2024housing}; \citealp{Blickle2019}; \citealp{benetton2022dynastic}), I replicate this result to highlight the importance of transfers for housing outcomes. I estimate
\begin{equation}	\label{eq:transferrent}
	\Pr\left(Own_{i,t+2}=1 \mid Own_{i,t}=0\right)
	= \alpha + \beta\,Transfer_{i,t+2}
	  + \gamma' X_{i,t} + \varepsilon_{i,t},
\end{equation}
where \(Own_{i,t}=0\) denotes a renter at time \(t\), \(Transfer_{i,t}=1\) if the household received a transfer between survey waves, and \(X_{i,t}\) includes log parental wealth, log household income and log household net worth, log household size, age, parental age, and dummies for  education, marital status, race, household composition changes, and state and year fixed effects. The sample is limited to renting households aged 25 to 44 who have never been observed as homeowners. Table \ref{tab:newowners_short} reports the estimated coefficient on transfer ($\beta$), with the full coefficient table relegated to Appendix Table \ref{tab:newowners}. These regressions are similar to those in \cite{Lee2018}, who also use the PSID.  It is important to note that transfer receipt is not exogenous---many transfers may be made precisely because the renter is buying a home.

I use two definitions of transfer receipt. The narrow definition includes parental gifts over \$500 in 2012, limiting the sample to renters in 2011. The broad definition includes any gift or inheritance over \$10,000 reported in any wave, or a large parental gift in 2012. This allows for household fixed effects and a larger sample, but captures only large transfers and includes some non-parental ones.

@@

To increase statistical power, I now use the broader transfer definition, which includes all large transfers. This expands the sample size tenfold but reduces the transfer receipt rate to around 3\%, while the median transfer size rises from \$1,500 to \$24,200. The broader sample also allows me to follow renters over time and include household fixed effects. In a univariate regression, transfer receipt is associated with a 21.9 percentage point increase in the probability of becoming a homeowner. Even with a full set of controls, the effect remains large and statistically significant: 13.5 percentage points in the OLS specification ($p<0.001$) and 7.9 percentage points with household fixed effects ($p<0.1$). These magnitudes are substantial compared to the baseline rent-to-own transition rate of 18\%.

@@

\subsubsection{Transfer Sizes}
For additional detail on the transfer variables, I plot kernel density estimates in Figure \ref{fig:PSID_transfer_to_renters}, for renters in $t$, separated by whether they own or rent in $t+2$, for the two transfer definitions. There are two takeaways. First, households who transition to ownership receive larger transfers on average, under both transfer measures. Second, the transfers in the transfer supplement are much smaller than than those in the gift/inheritance question. Indeed, the median parental transfer is about \$1k, the mean about \$3k, and the 95th percentile is at \$10k, compared to the median gift/inheritance of \$25k, mean of \$60k, and a 95th percentile of about \$130k.

\begin{figure}[tb]
	\caption{Transfer Sizes to Renters by Transition to Ownership}\label{fig:PSID_transfer_to_renters}
	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_transfers_renters_transfers.pdf}% 
	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_transfers_renters_inheritance.pdf}%

@@

While the PSID allows us to observe transfers and changes in housing tenure over time, it has several limitations. First, the transfer supplement captures transfers only in 2012, while the PSID is collected biennially, leading to undercounts of both transfer rates and amounts. Second, in the gift/inheritance question, the source of the transfer is unknown. Third, it is unclear whether intra-family loans are consistently reported or repaid.

\subsubsection{Households with Wealthier Parents Retain Ownership}
While most research focuses on the transition into ownership, I now focus on the challenge of retaining it.

\begin{figure}[tb]
	\captionsetup[subfigure]{aboveskip=-1pt}
	\caption{Retaining Homeownership}\label{fig:homeowner_exit}
	\begin{subfigure}{0.33\textwidth}%
		\caption{Retaining Ownership}\label{fig:owner_hazard}%
		\includegraphics[width=\textwidth]{../tabfig/descr/PSID_ownerexit.pdf}% 
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.33\textwidth}%
		\caption{All Owners}\label{fig:coef_allowner}%
		\includegraphics[width=\textwidth]{../tabfig/descr/PSID_coefowner.pdf}% 
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.33\textwidth}%
		\caption{First-Time Owners}\label{fig:coef_ftowner}%
		\includegraphics[width=\textwidth]{../tabfig/descr/PSID_coefftowner.pdf}% 
	\end{subfigure}%

	\caption*{\footnotesize \textit{Notes:} The left panel shows the share of current owners who remain owners in future years. The center and right panels show the estimated coefficients from Eq.~\ref{eq:regmaintain}, with dashed lines indicating 90\% confidence intervals. See main text for details.}
\end{figure}

@@

To understand which variables predicts retainment of ownership, I estimate
\begin{equation}	\label{eq:regmaintain}
	\Pr\left(Own_{i,t+\Delta}=1|Own_{i,t}=1\right)
	= \beta\ln(NetWorth_{i,t}) + \zeta\ln(ParWealth_{i,t})
	  + \gamma' X_{i,t} + \varepsilon_{i,t},
\end{equation}
where $Own_{i,t+\Delta}=1$ denotes an owner at time $t+\Delta$, $\beta$ and $\zeta$ measures the effect of log household net worth and parental wealth, respectively  and $X_{i,t}$ is a vector of control variables. The only changes relative to Eq. \ref{eq:transferrent} is that the outcome variable conditions on current ownership and measures future ownership, that there are no controls for transfer receipt, and that I have taken the two variables of interest out of the vector of control variables. The sample is limited to households aged 25 to 44. These regressions are motivated by \cite{bond2021role}, who estimate similar equations in the PSID and the Health and Retirement Study. Figures \ref{fig:coef_allowner} and \ref{fig:coef_ftowner} plots the two coefficients of interest over different time horizons.

Two results stand out. First, parental wealth is at least as predictive as household net worth at all horizons, although the difference is never statistically significant. Second, the effect of current parental wealth grows with the time horizon. For instance, a 10\% increase in parental wealth raises the probability of retaining ownership after eight years by 0.45 percentage points among first-time owners. I return to these results later as part of the model validation.

\subsubsection{Households with Wealthier Parents Are Less Delinquent} 
One potential roadblock to maintaining ownership is mortgage delinquency, as missed payments can lead to distressed or forced sales. I show that households with wealthier parents are less likely to fall behind on their mortgages, despite purchasing more expensive homes. The PSID has collected mortgage delinquency data since 2009; about 2\% of owners aged 25 to 44 report being behind.

To examine what predicts mortgage delinquency, I estimate: 
\begin{equation*}
\Pr(Behind_{i,t+2}) = \beta\ln(NetWorth_{i,t}) + \zeta\ln(ParWealth_{i,t}) + \gamma' X_{i,t} + \varepsilon_{i,t},
\end{equation*}
using the same control variables as in the previous regression. Table~\ref{tab:hypo} reports the estimated coefficients on household net worth ($\beta$); full results appear in Appendix Table~\ref{tab:hypo_long}.

I first limit the sample to first-time buyers to focus on distress among new owners. In a univariate regression, a 10\% increase in parental wealth is associated with a 0.07 percentage point decrease in the probability of being behind on mortgages two years later ($p<0.1$) With the full set of controls, the effect increases slightly, though it is no longer statistically significant. These magnitudes are economically meaningful since 1.6\% of first-time owners are behind on their mortgage.In Columns 3 to 5, I expand the analysis to include all owners who are currently not behind, increasing the sample sizes and allowing us to include household fixed effects. In these specifications, a 10\% increase in parental wealth is associated with a 0.09 percentage point decrease in delinquency risk in the univariate OLS regression ($p<0.001$), a 0.02 percentage point decrease in the multivariate OLS regression (not significant), and a 0.07 percentage point decrease in the fixed effects regression ($p<0.1$).

\begin{table}
	\centering
	\begin{threeparttable}
		\caption{Housing Choices and Parental Wealth}
		\label{tab:hypo}
		\small 
				\input{../tabfig/regr/hypoII_short.tex}

@@

Overall, the probability of falling behind on mortgage payments declines with parental wealth, even after controlling for a wide range of household characteristics. This highlights a novel channel through which parental resources help new homeowners retain ownership after purchase.

\subsubsection{Households with Wealthier Parents Are Less Likely to Downsize during Unemployment Spells}\label{sec:eventstudy}
Distressed and forced home sales are costly and often triggered by unemployment \citep{kermani2021racial,hsu2018unemployment}. To test whether parental wealth mitigates this risk, I perform a simple event study of how unemployment affects the likelihood of downsizing.

The exercise follows \cite{Chetty2007} closely. The outcome of interest is the change in log housing consumption, which is set to zero for households who do not move. For movers, housing consumption is defined as yearly rent while renting and as imputed rent for owners, set to 5\% of the market value \citep{Davis2008}. The sample is limited to household heads who experience only one unemployement spell between ages 25 and 44. I divide the sample by whether a household's parents were in the top parental wealth quartile at the time of unemployment and compare the two groups' housing consumption growth at unemployment. The results are displayed in Figure \ref{fig:housinggrowthrates}. Among households with non-wealthy parents, housing consumption falls by about 5\% during unemployment ($p<0.05$). In contrast, households with wealthy parents see a slight increase (not significant)in housing consumption during unemployment.\footnote{In Appendix \ref{app:eventstudy_details}, I also report the results of a formal event study using the same set of controls as in the previous regressions.}  I return to these results later as part of the model validation.

\begin{figure}
	\caption{Event Study: Housing Consumption at Unemployment by Parental Wealth}\label{fig:housinggrowthrates}
	\noindent\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthpoor_both}\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthrich_both}
